Correcting K-correction: dark energy is based on a math error from 1930

### 2025-01-10 paper/dark_energy_corrections.tex
More revisions.

\begin{abstract}
We rederive the K-correction equation and discover that the version commonly
used to calculate magnitudes contains three errors, and the oldest of these errors
traces back nine decades. The 1930 the equation did not correct for spectral
bandwidth stretching. Two new errors were introduced in 1996, one of which
causes systematic distance biases by inverting the zero-point correction term.
We conclude that properly correcting apparent magnitudes for redshift resolves
two of the preeminent mysteries in cosmology: first, dark energy is not
supported by observations of Type Ia supernovae and second, the Hubble tension
is due to the errors in the K-correction equation.
\end{abstract}

@@

The estimate of Type Ia supernovae (SnIa) distances is one of the primary
sources of data for cosmological models. Performing these distance estimates
involves taking a measurement of redshifted light in our observation frame and
then calculating the rest frame magnitude as if no redshift had occurred. It is
standard to report rest frame magnitudes in the $B$ filter, which is sensitive to
blue light, regardless of the filter used to take measurements in the
observation frame. The equation that converts from an observation frame
magnitude to a rest frame magnitude in the $B$ filter is called K-correction.

\citet{tolman1930} published the first formal derivation of K-correction, but
he did not account for spectral bandwidth stretching, one of the dimming
effects of redshift. An alternative K-correction was correctly derived by
\citet{desitter1934}. However, in \citet{hubble1935}, the discrepancy was noted
yet dismissed. \citet{oke1968} rederived the K-correction equation where they
accounted for spectral bandwidth stretching, but they did not account for time
dilation, one of the other dimming effects of redshift. The paper by
\citet{kim1996}, which is the basis for modern implementations of K-correction,
extended the work of \citet{oke1968} to handle additional cross-filter
comparisons and added a term that was intended to address zero-point
corrections. This history is explored in more depth in Section
\ref{sec:history}.

The effect of using an incorrect redshift correction for SnIa is that we think
objects are farther away than they really are; this effect compounds for
greater distances. Up until \citet{riess1998}, we did not have distant enough
observations for this error to matter much for cosmological models, but by the
late 1990s, it became clear that our measurements for distance and redshift
were not linear. This realization led to a model that utilized a cosmological
constant $\Lambda$ and dark energy in order to explain the non-linear
distance-redshift graph. We will show in Section \ref{sec:conclusions} that
correcting the flaws with K-correction leads to a linear graph that does not
need to rely on a cosmological constant.

@@

\section{A brief history of K-correction}
\label{sec:history}

@@

\begin{figure}
  \plotone{lum_distance_vs_redshift.pdf}
  \caption{The relationship between luminosity distance and redshift for two
  treatments of magnitude data. The displayed points are roughly a third of the
  values in the full SnIa dataset published in \citet{abbott2024}, selected
  with uniform redshift spacing to aid visibility. The red $k(z) = 1$ treatment,
  which represents the uncorrected values, is clearly non-linear while the blue
  $k(z) = 1 + z$ treatment, which fixes the K-correction redshift error,
  appears to be linear.}
  \label{fig:mu_distance_vs_redshift}
\end{figure}

SnIa are used to explore the relationship between distance and redshift because
these events always happen in similar ways, which keeps the absolute brightness
roughly constant. An analogy is to imagine someone walking in the dark and
lighting matches. If we know how brightly a match burns at a known distance, we
can estimate the distance to any match by measuring the apparent brightness
before applying some geometry.

@@

A few years later, \citet{desitter1934}, discussed all three issues that reduce
the observed magnitude of a distant observation. The correction for each of
these issues is identical: take a measurement for luminosity and multiply it by
the factor $1 + z$.

@@

The modern treatment of K-correction is based on the work of \citet{kim1996}.
This work extended the calculations of K-correction to extend to filters beyond
the blue $B$ and visible $V$ bands. It also introduced a term that deals with
the zero-point for the actual filters. Historically, bolometric devices would
measure the energy flux, but modern charge-coupled device (CCD) cameras
effectively measure the photon flux.

@@

With modern CCD cameras, a telescope observation consists of a single value
$\mathcal{F}_y$ erg/s, which represents the energy collected per second in
filter $y$. A summary of how this works is provided by \citet{lesser2015}.
Photogenerated electrons are collected in potential wells, which means that the
measured energy is proportional to the number of photons. However, we need to
calculate the expected number of photons collected by using a spectral energy
density function $F$. The value $F(\lambda)$ gives the amount of energy
collected by a bolometric device for the wavelength $\lambda$.

\begin{figure}
  \plotone{k_equation_flow.pdf}
  \caption{The value $K'_{xy}$ is equal to the number of photons we want to
  report in the target filter divided by the number of photons we measure in
  the observation filter.  Deriving this value starts with $F(\lambda)$, the
  spectral energy density, which we then convert to $F'(\lambda)$, the photon
  density for the rest frame.  Starting at the top right corner of the figure,
  we show the process of computing $R'(\lambda)$, the photon density in the
  observation frame, in three phases. First we stretch the photon density, but
  this step inflates counts. In the second step, we correct for the inflated
  counts by multiplying by $1/(1+z)$; at this point, the total count equals the
  total count of $F'(\lambda)$. Finally, in the third step we finish
  calculating the observation frame photon density $R'(\lambda)$ by accounting
  for time dilation, which also reduces photon counts by a factor of $1/(1+z)$.
  We can compute the expected ratio of photons by totalling the area under the
  curves in the bottom two panels.
  }
  \label{fig:k-example}
\end{figure}

@@

\noindent However, this equation is misleading and error prone. We will want to
use it to help calculate the amount of flux in an observation filter at the
redshifted wavelength $\lambda \times (1 + z)$. In other words, we want to
produce the photon density function $R'$ that is the redshifted version of
$F'$. Redshifting the photon density does two things:

@@

  In order to account for this spectral bandwidth stretching effect, we
  multiply $F'$ by $1/(1+z)$.

@@

In order to calculate the amount of flux $\mathcal{F}_x$ measured in filter
$x$, we need to compute the photon density $F'(\lambda)$ and multiply it by
the sensitivity $S_x(\lambda)$, which represents the proportion of photons with
wavelength $\lambda$ that filter $x$ will measure. We then need to sum over
all wavelengths, which is expressed with the equation

@@

We will also use the energy flux $\mathcal{F}$ to magnitude $m$ formula:

@@

It is useful to consider whether the term $P'_y / P'_x$ is correct or if
it should be inverted, but if we rewrite Equation \ref{eq:as_flux} as

\begin{equation}
  \mathcal{F}_x \times P'_x = \mathcal{F}_y \times P'_y \times K'_{xy}
\end{equation}

\noindent then the meaning becomes more clear. The left side, $\mathcal{F}_x
\times P'_x$, is the rate of photon collection for an idealized telescope in
filter $x$ while $\mathcal{F}_y \times P'_y$ is the rate of photon collection
for an idealized telescope in filter $y$. We observe the value $\mathcal{F}_y
\times P'_y$ and want to use the fudge factor $K'_{xy}$ to produce the value
$\mathcal{F}_x \times P'_x$.

We can isolate $K'_{xy}$ and then use Equation \ref{eq:flux_definition} to expand

@@

We now use Equations \ref{eq:redshifted_density} and \ref{eq:flux_definition}
to calculate the fluxes $\mathcal{F}_x$ and $\mathcal{F}_y$ in terms of the
spectral energy density function $F$. Note that $\mathcal{F}_x$ uses the rest
frame spectral energy density $F$, while $\mathcal{F}_y$ uses the redshifted
spectral energy density $R(\lambda)$:

\begin{equation}
\begin{aligned}
  K'_{xy} &= \frac{P'_x}{P'_y} \times \frac{\mathcal{F}_x}{\mathcal{F}_y} \\
         &= \frac{P'_x}{P'_y} \times
              \frac{\int F'(\lambda) S_x(\lambda) d\lambda}
                   {\int R'(\lambda) S_y(\lambda) d\lambda} \\
         &= \frac{P'_x}{P'_y} \times
              \frac{\int F'(\lambda) S_x(\lambda) d\lambda}
                   {\int F'(\lambda / (1+z)) \times (1 / (1 + z)^2) S_y(\lambda) d\lambda} \\
         &= \frac{P'_x}{P'_y} \times (1+z)^2 \times
              \frac{\int (\lambda / hc) F(\lambda) S_x(\lambda) d\lambda}
                   {\int (\lambda / ((1+z)hc)) F(\lambda / (1+z)) S_y(\lambda) d\lambda} \\
         &= \frac{P'_x}{P'_y} \times (1 + z)^2 \times
              \frac{\int \lambda F(\lambda) S_x(\lambda) d\lambda}
                   {\int (\lambda / (1+z) F(\lambda / (1+z)) S_y(\lambda) d\lambda} .
\end{aligned}
\end{equation}

Finally, we can use Equation \ref{eq:k_substitution} to convert $K'_{xy}$ back
into the magnitude $K_{xy}$:

@@

In order to fully compare our derivation against Equation \ref{eq:kim}, we need
to use the identity

@@

\begin{equation}
\begin{aligned}
  K_{xy} &= 5 \text{log} (1 + z)
            + 2.5 \text{log} \left(
              \frac{\int \lambda F(\lambda) S_x(\lambda) d\lambda}
                   {\int (\lambda / (1+z)) F(\lambda / (1+z)) S_y(\lambda) d\lambda} \right)
            - P_x + P_y \\
         &= 5 \text{log} (1 + z)
            + 2.5 \text{log} \left(
              \frac{\int \lambda F(\lambda) S_x(\lambda) d\lambda}
                   {\int (\lambda / (1+z)) F(\lambda / (1+z)) S_y(\lambda) d\lambda} \right) \\ &\
            + 2.5 \text{log} \left( \int (\lambda / hc) \mathcal{Z}(\lambda) S_x(\lambda) d\lambda \right)
            - 2.5 \text{log} \left( \int (\lambda / hc) \mathcal{Z}(\lambda) S_y(\lambda) d\lambda \right) \\
         &= 2.5 \text{log} \left(
              \frac{\int \lambda \mathcal{Z}(\lambda) S_x(\lambda) d\lambda}
                   {\int \lambda \mathcal{Z}(\lambda) S_y(\lambda) d\lambda}
             \right)
            + 5 \text{log} (1 + z)
            + 2.5 \text{log} \left(
              \frac{\int \lambda F(\lambda) S_x(\lambda) d\lambda}
                   {\int (\lambda / (1+z)) F(\lambda / (1+z)) S_y(\lambda) d\lambda} \right) .
\end{aligned}
\end{equation}

@@

\begin{itemize}
  \item The sign is different for the first term. This will have a subtle but
  frustrating effect for real measurements. The zero-point values are usually,
  but not always, close to zero. As we observe more distant supernovae, it is
  useful to switch our observational filter to a filter that is sensitive to
  longer wavelengths. Each observation filter will have its own small bias.

  This effect by itself can alter estimates of $H_0$ by several percent. While
  we present an estimate of the Hubble constant in Figure
  \ref{fig:H0bootstrap}, we are not terribly confident in the numbers because
  we were unable to correct for the zero-point errors, or verify that these
  errors may not exist.

  \item The second term is multiplied by $2.5$ in \citet{kim1996}, but is
  multiplied by $5$ here. This is the manifestation of the error in
  \citet{tolman1930}.

  \item In our derivation, the third term has an extra $1 + z$ expression.
  Based on an inspection of the SN(oo)py software package presented by
  \citet{snpy}, this error is ignored and software package authors implement it
  as intended, not precisely as written.
\end{itemize}

@@

As shown in Figure \ref{fig:mu_distance_vs_redshift}, when reported magnitudes
are corrected by adding the magnitude $-2.5\text{log}(1+z)$, there is a linear
relationship between redshift and luminosity distance. This is in accordance
with the Hubble-Lema\^{i}tre law. As opposed to the uncorrected values, there
is no visual acceleration. Since the corrected values are approximately linear,
we can use them to estimate the Hubble constant $H_0$. This is demonstrated in
Figure \ref{fig:expansion}. Furthermore, we performed a bootstrap calculation
to estimate the distribution of $H_0$ estimates in Figure
\ref{fig:H0bootstrap}.

\begin{figure}
  \plotone{bootstrapped_H0.pdf}
  \caption{A histogram of $10{,}000$ bootstrap trials measuring the Hubble constant
  $H_0$. Each trial samples an absolute magnitude $M \sim Norm(-19.2334,
  0.0404)$ for SnIa based on data published by \citet{camarena2020}. It then samples,
  with replacement, a population of supernovae from the dataset published by
  \citet{abbott2024}. Finally, it uses the non-parametric linear regression
  technique described by \citet{siegel1982}. The result of the bootstrap is $H_0
  \sim Norm(65.94, 1.3^2)$.
  }
\label{fig:H0bootstrap}
\end{figure}

The value estimated here, $H_0 = 65.94 \pm 1.3$ is highly suspicious
because of the flipped zero-point corrections, but it is consistent with
estimates of $H_0$ that are based on the CMB. \citet{planck2020} published the
CMB value of $H_0 = 67.27 \pm 0.6$, and the one sigma error bars overlap. It
is interesting to note that restricting the Dark Energy Survey observations of
SnIa \citep{vincenzi2024} to only those with $z < 0.1$, which presumably is a
small enough redshift that all observations will have been made with the same
filter, gives the estimate $H_0 = 67.8 \pm 1.9$.

>>>>

\begin{abstract}
We rederive the K-correction equation and discover that the version commonly
used to calculate magnitudes contains multiple errors. The oldest of these
errors traces back to 1930 and two new errors were introduced in 1996. We
conclude that properly correcting apparent magnitudes for redshift resolves two
of the preeminent mysteries in cosmology: first, dark energy is not supported
by observations of Type Ia supernovae and second, the Hubble tension is due to
the errors in the K-correction equation.
\end{abstract}

@@

The estimate of Type Ia supernovae (SnIa) distances is one of the primary
sources of data for cosmological models. Performing these distance estimates
involves taking a measurement of redshifted light in our observation frame and
then calculating the rest frame magnitude as if no redshift had occurred. The
equation that converts from an observation frame magnitude to a rest frame
magnitude is called K-correction.

\citet{tolman1930} published the first formal derivation of K-correction, but
he did not account for spectral bandwidth stretching, one of the dimming
effects of redshift. An alternative K-correction was correctly derived by
\citet{desitter1934}. However, in \citet{hubble1935}, the discrepancy was
briefly discussed yet dismissed. \citet{oke1968} rederived the K-correction
equation where they accounted for spectral bandwidth stretching, but they did
not account for time dilation, one of the other dimming effects of redshift.
The paper by \citet{kim1996}, which is the basis for modern implementations of
K-correction, extended the work of \citet{oke1968} to handle additional
cross-filter comparisons and added a term that was intended to address
zero-point corrections. This history is explored in more depth in Section
\ref{sec:history}.

The effect of using an incorrect redshift correction for SnIa is that we think
objects are farther away than they really are; this effect compounds for
greater distances. Up until \citet{riess1998}, we did not have distant enough
observations for this error to matter much for cosmological models, but by the
late 1990s, it became clear that our measurements for distance and redshift
were not linear. This realization led to a model that utilized a cosmological
constant $\Lambda$ and dark energy in order to explain the non-linear
distance-redshift graph. We will show in Section \ref{sec:conclusions} that
correcting the flaws of K-correction leads to a linear graph that does not
need to rely on a cosmological constant.

@@

\section{
  A brief history of K - correction
}
\label{sec:history}

@@

\begin{figure}
  \plotone{lum_distance_vs_redshift.pdf}
  \caption{The relationship between luminosity distance and redshift for two
  treatments of magnitude data. The red $k(z) = 1$ treatment, which represents
  the uncorrected values, is clearly non-linear while the blue $k(z) = 1 + z$
  treatment, which fixes the K-correction redshift error, appears to be linear.
  The displayed points are roughly a third of the values in the full SnIa
  dataset published in \citet{abbott2024}, selected with uniform redshift
  spacing to aid visibility. }
  \label{fig:mu_distance_vs_redshift}
\end{figure}

SnIa are used to explore the relationship between distance and redshift because
these events have a roughly constant absolute magnitude, which allows us to use
the apparent magnitude to estimate distance. An analogy is to imagine someone
walking in the dark and lighting matches. If we know how brightly a match burns
at a known distance, we can estimate the distance to any match by measuring the
apparent brightness before applying some geometry.

@@

A few years later, \citet{desitter1934} discussed all three issues that reduce
the observed magnitude of a distant observation. The correction for each of
these issues is identical: take a measurement for luminosity and multiply it by
the factor $1 + z$.

@@

The modern treatment of K-correction is based on the work of \citet{kim1996}.
This work extended the work of \citet{oke1968} to apply to filters beyond the
blue $B$ and visible $V$ bands. It also introduced a term that deals with the
zero-point for the actual filters. Historically, bolometric devices would
measure the energy flux, but modern charge-coupled device (CCD) cameras
effectively measure the photon flux, so \citet{kim1996} derived a form of the
K-correction equation that uses photon flux.

@@

The K-correction $K_{xy}$ uses a spectral energy density $F$ template that
specifies the expected spectrum for a SnIa observation. It also uses the
observed redshift $z$ and the observed photon flux magnitude $m_y$.

With modern CCD cameras, a telescope observation consists of a single value
$\mathcal{F}_y$ erg/s, which represents the energy collected per second in
filter $y$. Photogenerated electrons are collected in potential wells, which
means that the energy measured in this filter is proportional to the number of
photons \citep{lesser2015}. However, we need to calculate the expected number
of photons collected by using a spectral energy density function $F$.

\begin{figure}
  \plotone{k_equation_flow.pdf}
  \caption{The value $K'_{xy}$ is equal to the number of photons we want to
  report in the target filter divided by the number of photons we measure in
  the observation filter. Deriving this value starts with $F(\lambda)$, the
  spectral energy density, which we then convert to $F'(\lambda)$, the photon
  density for the rest frame. Starting at the top right corner of the figure we
  show three steps used to compute $R'(\lambda)$, the photon density in the
  observation frame. We first stretch the photon density, but this step
  inflates counts. In the second step, we correct for the inflated counts by
  multiplying by $1/(1+z)$; at this point, the total count equals the total
  count of $F'(\lambda)$. Finally, in the third step we finish calculating the
  observation frame photon density $R'(\lambda)$ by accounting for time
  dilation, which also reduces photon counts by a factor of $1/(1+z)$. We can
  compute the expected ratio of photons by totalling the area under the curves
  in the bottom two panels.
  }
  \label{fig:k-example}
\end{figure}

@@

\noindent However, this equation is misleading and error prone. We will want to
use it to help calculate the flux in an observation filter at the redshifted
wavelength $\lambda \times (1 + z)$. In other words, we want to produce the
photon density function $R'$ that is the redshifted version of $F'$.
Redshifting the photon density does two things:

@@

  In order to account for the spectral bandwidth stretching effect, we
  multiply $F'$ by $1/(1+z)$.

@@

In order to calculate the flux $\mathcal{F}_x$ measured in filter $x$, we need
to compute the photon density $F'(\lambda)$ and multiply it by the sensitivity
$S_x(\lambda)$, which represents the proportion of photons with wavelength
$\lambda$ that filter $x$ will measure. We then sum over all wavelengths, which
is expressed with the equation

@@

We will also use the flux $\mathcal{F}$ to magnitude $m$ formula:

@@

\noindent We can isolate $K'_{xy}$ and then use Equation \ref{eq:flux_definition} to expand

@@

\noindent We now use Equations \ref{eq:redshifted_density} and \ref{eq:flux_definition}
to calculate the fluxes $\mathcal{F}_x$ and $\mathcal{F}_y$ in terms of the
spectral energy density function $F$. Note that $\mathcal{F}_x$ uses the rest
frame spectral energy density $F$, while $\mathcal{F}_y$ uses the redshifted
spectral energy density $R(\lambda)$:

\begin{equation}
\begin{aligned}
  K'_{xy} &= \frac{P'_x}{P'_y} \times \frac{\mathcal{F}_x}{\mathcal{F}_y} \\
         &= \frac{P'_x}{P'_y} \times
              \frac{\int F'(\lambda) S_x(\lambda) d\lambda}
                   {\int R'(\lambda) S_y(\lambda) d\lambda} \\
         &= \frac{P'_x}{P'_y} \times
              \frac{\int F'(\lambda) S_x(\lambda) d\lambda}
                   {\int F'(\lambda / (1+z)) \times (1 / (1 + z)^2) S_y(\lambda) d\lambda} \\
         &= \frac{P'_x}{P'_y} \times (1+z)^2 \times
              \frac{\int (\lambda / hc) F(\lambda) S_x(\lambda) d\lambda}
                   {\int (\lambda / ((1+z)hc)) F(\lambda / (1+z)) S_y(\lambda) d\lambda} \\
         &= \frac{P'_x}{P'_y} \times (1 + z)^2 \times
              \frac{\int \lambda F(\lambda) S_x(\lambda) d\lambda}
                   {\int (\lambda / (1+z)) F(\lambda / (1+z)) S_y(\lambda) d\lambda} .
\end{aligned}
\end{equation}

Finally, we use Equation \ref{eq:k_substitution} to convert $K'_{xy}$ back into
the magnitude $K_{xy}$:

@@

In order to fully compare our derivation against Equation \ref{eq:kim} (from
\citet{kim1996}) we need to use the identity

@@

\begin{equation}
\begin{aligned}
  K_{xy} &= 5 \text{log} (1 + z)
            + 2.5 \text{log} \left(
              \frac{\int \lambda F(\lambda) S_x(\lambda) d\lambda}
                   {\int (\lambda / (1+z)) F(\lambda / (1+z)) S_y(\lambda) d\lambda} \right)
            - P_x + P_y \\
         &= 5 \text{log} (1 + z)
            + 2.5 \text{log} \left(
              \frac{\int \lambda F(\lambda) S_x(\lambda) d\lambda}
                   {\int (\lambda / (1+z)) F(\lambda / (1+z)) S_y(\lambda) d\lambda} \right) \\ &\ \ \
            + 2.5 \text{log} \left( \int (\lambda / hc) \mathcal{Z}(\lambda) S_x(\lambda) d\lambda \right)
            - 2.5 \text{log} \left( \int (\lambda / hc) \mathcal{Z}(\lambda) S_y(\lambda) d\lambda \right) \\
         &= 2.5 \text{log} \left(
              \frac{\int \lambda \mathcal{Z}(\lambda) S_x(\lambda) d\lambda}
                   {\int \lambda \mathcal{Z}(\lambda) S_y(\lambda) d\lambda}
             \right)
            + 5 \text{log} (1 + z)
            + 2.5 \text{log} \left(
              \frac{\int \lambda F(\lambda) S_x(\lambda) d\lambda}
                   {\int (\lambda / (1+z)) F(\lambda / (1+z)) S_y(\lambda) d\lambda} \right) .
\end{aligned}
\end{equation}

@@

\begin{itemize}
  \item The sign is different for the first term. This will have a subtle but
  frustrating effect for real measurements. The zero-point values for the two
  filters are usually, but not always, roughly equal. As we observe more
  distant supernovae, it is useful to switch our observational filter to a
  filter that is sensitive to longer wavelengths, but each observation filter will
  have its own small bias.

  This effect by itself can alter estimates of $H_0$ by several percent. While
  we present an estimate of the Hubble constant in Figure
  \ref{fig:H0bootstrap}, we are not terribly confident in the numbers because
  we were unable to correct the zero-point errors, or verify that these errors
  may not exist.

  \item The second term is multiplied by $2.5$ in \citet{kim1996}, but is
  multiplied by $5$ here. This is the manifestation of the error in
  \citet{tolman1930} and leads to the conclusion that the expansion of the
  universe is accelerating.

  \item In our derivation, the third term has an extra $1 + z$ expression.
  Based on an inspection of the SN(oo)py software package presented by
  \citet{burns2010} and the SNANA software package presented by
  \citet{kessler2009}, this error is ignored and software package authors
  implement it as intended, not precisely as written.
\end{itemize}

@@

As shown in Figure \ref{fig:mu_distance_vs_redshift}, when reported magnitudes
are corrected by adding the magnitude $-2.5\text{log}(1+z)$, there is a linear
relationship between redshift and luminosity distance. This is in accordance
with the Hubble-Lema\^{i}tre law. As opposed to the uncorrected values, there
is no visual acceleration. Since the corrected values are approximately linear,
we can use them to estimate the Hubble constant $H_0$ which is demonstrated in
Figure \ref{fig:expansion}. Furthermore, we performed a bootstrap calculation
to estimate the distribution of $H_0$ estimates in Figure
\ref{fig:H0bootstrap}.

\begin{figure}
  \plotone{bootstrapped_H0.pdf}
  \caption{A histogram of $10{,}000$ bootstrap trials measuring the Hubble constant
  $H_0$. Each trial samples an absolute magnitude $M \sim Norm(-19.2334,
  0.0404)$ \citep{camarena2020} before it samples, with replacement, a
  population of supernovae from the dataset published by \citet{abbott2024}.
  Finally, it uses the non-parametric linear regression technique described by
  \citet{siegel1982}. The result of the bootstrap is $H_0 \sim Norm(65.94, 1.3^2)$.
  }
\label{fig:H0bootstrap}
\end{figure}

The value estimated here, $H_0 = 65.94 \pm 1.3$ is highly suspicious
because of the flipped zero-point corrections, but it is consistent with
estimates of $H_0$ that are based on the CMB. \citet{planck2020} published the
CMB value of $H_0 = 67.27 \pm 0.6$, and the one sigma error bars overlap. It
is interesting to note that restricting the Dark Energy Survey observations of
SnIa \citep{vincenzi2024} to only those with $z < 0.1$, which presumably is a
small enough redshift that most of the observations will have been made with
the same filter, gives the estimate $H_0 = 67.8 \pm 1.9$.
